---
title: "Reproducible Research Project 1"
output: html_document
---
#Introduction
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


## Accessing data
```{r}
data1 <- read.csv("activity.csv")
data_sum <- aggregate(data1$steps, by=list(Category=data1$date), FUN=sum)
```

## Histogram of number of steps taken each day
The data is aggregated by date and a histogram produced displaying the data
```{r}
hist(data_sum$x, main="Total steps per day", xlab="Steps", ylab="Frequency")
```

## Mean and median number of steps taken each day
```{r}
data_mean1 <- mean(data1$steps, na.rm=TRUE)
print(paste("The mean steps per day is: ", data_mean1))

data_median1 <- median(data1$steps, na.rm=TRUE)
print(paste("The median steps per day is: ", data_median1))
```

## Time series plot of the average number of steps taken by interval
```{r}
data_mean_interval <- aggregate(data1$steps, by=list(Category=data1$interval), FUN=mean, na.rm=TRUE)
plot(data_mean_interval$Category, data_mean_interval$x, type="l", main="Average steps per five minute interval", xlab="Interval No", ylab="steps")
```

## 5 minute interval with max values
```{r}
maxsteps <- max(data_mean_interval$x)
print(paste("The maximum number of steps in a five minute interval was: ", maxsteps))
maxinterval <- data_mean_interval[data_mean_interval$x==max(data_mean_interval$x),]
print(paste("The interval with maximum number of steps was: ", maxinterval$Category))
```

## Number of rows with missing values 
There are a number of missing values in the dataset
```{r}
sum(!complete.cases(data1)) 
```

## Strategy for imputing missing values - Use the median
I have chosen to replace missing values with median
```{r}
data2 <- data1
data2$steps[is.na(data2$steps)] <- median(data2$steps,na.rm=TRUE)
sum(!complete.cases(data2)) 
data_sum_impute <- aggregate(data1$steps, by=list(Category=data1$date), FUN=sum)
hist(data_sum_impute$x)
```


## Calculating new mean and median
```{r}
data_mean2 <- mean(data2$steps)
print(paste("The mean is: ", data_mean2))
data_median2 <- median(data2$steps)
print(paste("The median is: ", data_median2))
```
Imputing missing values lowers the mean but has no effect on median

## Differences in activity patterns between weekdays and weekends
To determine if there are activity differences between weekends and weekdays, I have plotted two graphs - one with weekday data and one with weekend data. This lets us easily see the differences in the pattern between weekday and weekend
```{r}
data2$date <- as.Date(data2$date)
data2$dayname <- weekdays(data2$date)
data2$weekend <- as.factor(ifelse(data2$dayname == "Saturday" |
                                    data2$dayname == "Sunday", "weekend", "weekday"))
library(lattice)
plotdata <- aggregate(steps ~ interval + weekend, data2, mean)
xyplot(steps ~ interval | factor(weekend), data=plotdata, aspect=1/3, type="l")
```


